Author:   Bibobra Alabrah
title:    United States of America Birth Records Analysis
Dataset:  US Present Birth Records
Date:     6/7/2018
---

Import dependencies(libraries)
```{r load-packages, message = FALSE}
library(dplyr)
library(ggplot2)
library(statsr)
```

Load the dataset:
```{r load-present-data}
data(present)
```

View the number of records and features of the dataset:
```{r dimensions-of-present}
dim(present)
```

Rename the dataset to birth_records:
```{r}
birth_records <- present
rm(present)
```

View the first 10 records of the dataset
```{r}
head(birth_records, 10)
```

What years are included in this dataset? 
```{r years-in-present-data}
range(birth_records$year)
```
We see that the birth records span from 1940 to 2013.

What is the total number of births for each year?
```{r}
birth_records <- birth_records %>%
  mutate(total = boys + girls)
head(birth_records)
```

What is the proportion of boys born each year?
```{r}
birth_records <- birth_records %>%
  mutate(prop_boys = boys/total)
head(birth_records$prop_boys)
```

What about the proportion of girls?
```{r}
birth_records <- birth_records %>%
  mutate(prop_girls = girls/total)
head(birth_records$prop_girls)
```
We can see that generally more boys were born during this time.
Let us visualize this information.

```{r}
ggplot(birth_records, aes(x = prop_boys, y = prop_girls)) +
  geom_point()+
  geom_line()
```

Let us Plot these values over time and based on the plot determine if the 
following statement is true or false: The proportion of boys born in the US has 
decreased over time. 

```{r prop-boys-over-time}
ggplot(birth_records, aes(x = year, y = prop_boys)) +
  geom_point()+
  geom_line()
```
Based on the plot we can see that the proportion of boys born in the US has 
decreased over time.

Has the number of girls increased over time?
```{r}
ggplot(birth_records, aes(x = year, y = prop_girls)) +
  geom_point()+
  geom_line()
```
Based on the information, the number of girls has increased over time.

What is the boy to girl ratio for each?
```{r}
birth_records <- birth_records %>%
  mutate(prop_boy_girl = boys/girls)
```

Plot these values over time. Describe the observed trend? 
```{r}
ggplot(birth_records, aes(x = year, y = prop_boy_girl)) +
  geom_point()
```
There is initially a decrease in the boy-to-girl ratio, and then an increase between 1960 and 1970, followed by a decrease. 

8. In what year did we see the most total number of births in the U.S.? 

```{r most-total-births}
birth_records %>%
 mutate(total = total) %>%
arrange(desc(total))
```
We see that the US had the most total number of births in 2007.